\begin{abstract}
Do user-stated quality thresholds (``this only needs to be a 70 out of 100'') actually constrain LLM revision behavior? In a two-turn writing setup with 3,840 trials (Study~1), we cross follow-up prompt phrasing (2 levels), stated quality threshold (8 levels), and model (Claude Sonnet 4, Gemini 2.5 Flash, GPT-4o) across eight scenarios. Quality thresholds have \textit{no detectable effect} on whether the model revises. A revision-implying follow-up (``Can this be improved?'') triggers near-universal revision (99.9\%) regardless of threshold, while an evaluative follow-up produces 0.3--38.0\% revision rates depending on model. Within triggered revisions, thresholds produce only a weak-to-moderate dose-response (Spearman $\rho = -0.14$ to $-0.50$). These results suggest a two-stage pattern: prompt phrasing controls whether the model revises; the stated threshold modulates revision intensity only after revision is triggered. A follow-up momentum experiment (Study~2, 1,728 trials) tests whether prior revision rounds shift the gate on a subsequent evaluative probe. Revision history nearly doubles the overall revision rate ($\chi^2 = 376.54$, $p < 0.0001$), but the effect is almost entirely model-specific: GPT-4o jumps from 31\% to 98\% revision with a single prior round, while Claude Sonnet 4 and Gemini 2.5 Flash show no increase or slight decline. A reverse momentum condition (576 trials) shows the effect is bidirectional: a single round of affirming feedback (``This looks great, no changes needed'') eliminates full revision entirely and drops GPT-4o from 31\% to 1\%, confirming that the gate tracks the most recent conversational signal rather than the user's stated quality preferences.
\end{abstract}
