\section{Conclusion}

Across two studies and a reverse momentum condition totaling over 6,000 scored trials, we find that the phrasing of the follow-up prompt, not the stated quality threshold, determines whether the model revises. Models possess the ability to calibrate revision intensity to thresholds, but this ability is not activated when the follow-up prompt implies revision.

Study~2 extends this finding to multi-turn conversations, showing that prior revision history can shift the gate on a subsequent evaluative probe. This effect is sharply model-specific: GPT-4o's revision rate jumps from 31\% to 98\% with a single prior round, while Claude and Gemini resist. The reverse momentum condition shows that a single round of affirming feedback eliminates full revision entirely and suppresses GPT-4o from 31\% to 1\%, confirming that the gate tracks the most recent conversational signal in both directions. The threshold remains invisible to the gate regardless of momentum ($p = 0.51$).

For alignment research, these findings locate a threshold-blind compliance pattern at a specific decision point where the calibration ability models demonstrate goes unused. The model divergence under momentum suggests that this compliance pattern is not inherent to language modeling but is shaped by alignment training, and that different training strategies produce qualitatively different susceptibilities. Future work should test whether targeted interventions (e.g., injecting thresholds into the system prompt) can restore threshold sensitivity at the gate, and whether human evaluators converge with LLM judges on overcorrection severity.
