\section{Discussion}

\paragraph{The Two-Stage Account.}
\label{sec:discussion-two-stage}
The results presented above are consistent with a two-stage account of LLM revision behavior (Figure~\ref{fig:two-stage}) in which the decision to revise and the calibration of that revision respond to different input signals. Stage~1 (the revision gate) responds to the pragmatic force of the follow-up prompt with near-binary sensitivity: any phrasing that implies revision triggers compliance; any phrasing that implies evaluation suppresses it. Stage~2 (revision intensity) responds to the user's stated quality threshold with a weak-to-moderate dose-response. The probe $\times$ threshold interaction term is statistically significant ($\beta = -2.20$, $z = -5.69$, $p < 0.001$), but this is consistent with a floor effect rather than true coupling: because the evaluative probe suppresses revision almost entirely, there is no overcorrection variance for the threshold to modulate. Within each probe condition taken separately, the two stages appear to operate on separate inputs: the threshold has no detectable effect on the gate (Kruskal-Wallis $p > 0.40$), and probe phrasing does not modulate calibration intensity within the leading-probe condition.

This pattern is the central observation of the paper. It means that the calibration ability demonstrated by Stage~2, which shows models \textit{do} modulate revision intensity based on thresholds, is effectively inaccessible at Stage~1 under the conversational conditions we tested. Users who state quality preferences and then ask casual follow-up questions receive revisions that are insensitive to their stated standards, not because the model lacks calibration ability, but because the gate fires before calibration is reached.

\paragraph{Threshold-Blind Compliance.}
A natural objection is that ``Can this be improved?'' is a leading question with a pragmatically obvious answer: almost anything \textit{can} be improved. A model that answers ``yes'' and demonstrates an improvement may be following instructions appropriately, not behaving sycophantically. Whether the gate reflects sycophancy or straightforward instruction-following, the key observation is the same: the gate is threshold-blind. A well-calibrated instruction-follower would incorporate the user's stated quality bar: if the initial output meets a threshold of 70, and the user asks whether it can be improved, the model should be \textit{more likely to decline} than if the threshold were 100. The complete absence of threshold modulation at the gate (Kruskal-Wallis $p > 0.40$) shows that the model is not incorporating the user's stated preferences when deciding whether to revise. This pattern is consistent with the sycophancy literature \cite{perez2023discovering, sharma2024towards}, where models default to compliance when a prompt can be interpreted as a request, but we cannot distinguish sycophancy from threshold-insensitive instruction-following on the basis of behavioral data alone.

\paragraph{The Compliance Cliff.}
The five-probe pilot reveals that the transition between ``revise'' and ``don't revise'' is not a smooth gradient but a steep cliff. Probes that imply revision (``Can this be improved?'', ``Is there anything you would change?'') produce 100\% compliance; probes that imply evaluation (``What do you think?'', ``Review against the threshold'') produce 0--12.5\% revision rates. Only one probe fell in the intermediate range (25\%), suggesting that the cliff's edge is narrow. We note that five probes cannot definitively distinguish a binary cliff from a steep continuous function; finer-grained probe spacing is needed to characterize the transition zone. This binary sensitivity has practical implications: small changes in follow-up phrasing can dramatically alter model behavior, and users are unlikely to be aware of this sensitivity. The pattern is consistent with something closer to a speech-act classifier than a graded compliance function, in which the model detects whether a follow-up constitutes a directive and, if so, complies categorically, though we cannot confirm this from behavioral data alone.

\paragraph{Model Differences.}
The evaluative probe reveals strikingly different revision dispositions across the three models tested. Gemini 2.5 Flash declines revision in 99.7\% of evaluative-probe trials, while GPT-4o and Claude Sonnet 4 show more moderate decline rates (68.6\% and 62.0\%, respectively). Notably, Gemini shows the strongest overcorrection tendency when it does revise ($\beta = 1.15$ in Model~C) but is also the model most likely to correctly decline. This is consistent with Gemini having a more binary revision gate, while GPT-4o and Claude show more intermediate behavior between the evaluative and revision modes. This variation may reflect differences in fine-tuning procedures and instruction data composition \cite{ouyang2022training, bai2022training}, though we cannot identify the specific cause from behavioral data. All three models show comparable threshold sensitivity within the leading-probe condition, suggesting that the divergence is specific to the gate rather than to calibration ability. Understanding which training choices produce sharper or softer gates is an open question for alignment research.

\paragraph{Framing Effects.}
Qualitative threshold framings (``good enough to get the point across'') produce less overcorrection than numeric framings (``around a 70 out of 100''), with a consistent effect across all regression models ($\beta = -0.74$ to $-1.05$, all $p < 0.001$). One interpretation is that qualitative framings provide quality descriptions that are more grounded in the model's training distribution, while numeric scores may be harder for the model to map onto revision behavior. This hypothesis could be tested by providing models with explicit rubric mappings between numeric and qualitative thresholds and measuring whether the framing gap closes. Framing does not affect the revision gate, confirming that this effect operates at Stage~2.

\paragraph{Recency and Positional Effects.}
An alternative explanation for the gate's threshold insensitivity is that the threshold (embedded in Turn~1) is simply less salient than the probe (arriving in Turn~2). Transformer-based models are known to weight recent context more heavily in some settings. If the gate's behavior reflects recency bias rather than a genuine inability to incorporate thresholds, then injecting the threshold into the system prompt or repeating it in Turn~2 should restore sensitivity. We did not test this manipulation, but it represents a promising intervention and a way to distinguish the recency hypothesis from the two-stage account.

\paragraph{Design Implications.}
Our findings suggest that interfaces exposing user-stated quality preferences must also protect those preferences from being overridden by downstream conversational pragmatics. A system that allows users to set a quality bar but then routes follow-up messages through a standard chat interface effectively nullifies the quality constraint at the revision gate. These mitigations are speculative and untested, but follow directly from our findings: (1) injecting the quality threshold into the system prompt rather than the user message, (2) rewriting follow-up prompts to reference the stated threshold explicitly, or (3) adding a revision-gate classifier that checks whether the user's follow-up actually requests revision before passing it to the model.
