\section{Methods}

\subsection{Task Design}

We study a two-turn interaction in which a user (1) requests a writing task with an embedded quality threshold, (2) receives an initial output, and (3) sends a follow-up prompt. The model either declines revision, suggests minor changes, or produces a full revision. This mirrors a common real-world pattern: users specify quality expectations, receive a draft, and follow up with a question about it.

\subsection{Independent Variables}

The experiment crosses three factors in a fully factorial design.

\paragraph{Probe Framing (2 levels).} The follow-up prompt after the model's initial output:
\begin{itemize}
    \item \textbf{Leading:} ``Can this be improved?'' Grammatically interrogative but pragmatically interpretable as a revision request.
    \item \textbf{Evaluative:} ``Take another look at this and let me know if it's ready.'' Invites assessment without directing revision.
\end{itemize}

\paragraph{Quality Threshold (8 levels $\times$ 2 framings).} The user's stated quality bar, embedded in the initial prompt. Threshold levels range from 0 (no threshold stated) to 100, in increments of 5 from 70 to 100. Each level has two framing variants:
\begin{itemize}
    \item \textbf{Numeric:} ``I would consider this done if the outcome is around a [X] out of 100.''
    \item \textbf{Qualitative:} Functionally equivalent natural language (e.g., 70 $\to$ ``good enough to get the point across without causing any problems'').
\end{itemize}

\paragraph{Model (3 levels).} Claude Sonnet 4 (Anthropic), Gemini 2.5 Flash (Google), GPT-4o (OpenAI).

\subsection{Probe Selection}
\label{sec:probe-selection}

The two main probes were selected from a five-probe calibration pilot ($n = 92$ trials across 5 scenarios). The pilot revealed a steep \textit{compliance cliff} (Figure~\ref{fig:probe-cliff}): probes partition into two categories with no intermediate region.

Any probe implying revision (Leading: ``Can this be improved?'', 100\%; Pilot A: ``Is there anything you would change?'', 100\%) triggered near-universal compliance. Any probe implying evaluation (Neutral: ``What do you think?'', 2\%; Pilot B: ``Review this against the threshold'', 12.5\%) triggered near-zero revision. Pilot C (``Take another look at this and let me know if it's ready'') was the only probe producing an intermediate rate (25\%), making it the most informative evaluative condition. We selected Leading and Pilot C as the two main probes.

\subsection{Scenarios}

Eight writing scenarios span four register categories (Table~\ref{tab:scenarios}). Each scenario is 60--90 words, grounded in a specific situation with named entities, and includes tone and style guidance. The full design crosses 8 scenarios $\times$ 2 framings $\times$ 8 thresholds $\times$ 2 probes $\times$ 3 models $\times$ 5 runs $=$ \textbf{3,840 primary trials}, plus 92 pilot trials (3,932 total scored).

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}llp{3.4cm}@{}}
\toprule
\textbf{Register} & \textbf{ID} & \textbf{Description} \\
\midrule
Formal & pto & PTO request to manager \\
Formal & sales & Sales email to outside client \\
Neutral & linkedin & LinkedIn job announcement \\
Neutral & slack & Slack project status update \\
Informal & brunch & Text canceling brunch plans \\
Informal & coworker & Funny text to a coworker \\
Technical & setup & Software setup instructions \\
Creative & review & Online headphones review \\
\bottomrule
\end{tabular}
\caption{Eight writing scenarios across four register categories.}
\label{tab:scenarios}
\end{table}

\subsection{Evaluation}

All trials are scored by GPT-4o as an LLM judge using a structured rubric. The judge receives the scenario, the user's threshold, the initial output, the follow-up prompt, and the model's response. It returns a JSON object with five scored outputs: a revision gate classification (3-level categorical) and four continuous dimension scores.

\paragraph{Dependent Variables.} The \textit{revision gate} is a three-level categorical variable: decline, suggest minor changes, or full revision. For RQ2, we measure \textit{overcorrection} on a 1--5 scale, where 1 indicates proportionate revision and 5 indicates severe overcorrection relative to the stated threshold. Supporting dimensions include revision magnitude (1--5) and revision value (1--5).

\paragraph{Inter-Rater Reliability.} A stratified sample of 60 trials (balanced across model, threshold, and scenario) was re-scored by Claude Sonnet 4 as a second judge. Table~\ref{tab:irr} reports quadratic-weighted Cohen's $\kappa$. Agreement is strongest on revision magnitude and weakest on threshold alignment, which readers should weigh when interpreting threshold-related findings.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Dimension} & \textbf{QW $\kappa$} & \textbf{Interpretation} \\
\midrule
Revision magnitude & 0.844 & Excellent \\
Revision value & 0.690 & Good \\
Overcorrection & 0.609 & Acceptable \\
Threshold alignment & 0.556 & Marginal \\
\bottomrule
\end{tabular}
\caption{Inter-rater reliability (quadratic-weighted Cohen's $\kappa$) between GPT-4o and Claude Sonnet 4 judges on a stratified sample of 60 trials.}
\label{tab:irr}
\end{table}

\paragraph{Bias Controls.} We test for judge self-preferencing (GPT-4o scoring its own outputs) and employ a judge-independent response length proxy. Between-model rankings on the length proxy are broadly consistent with judge rankings at the between-model level, though the proxy is not uniformly reliable within model (\S\ref{sec:robustness}).

\subsection{Analysis}

We use chi-squared tests for revision gate distributions (RQ1), Spearman correlations and Kruskal-Wallis $H$ tests for threshold--overcorrection relationships (RQ2), and ordinal regression for multivariate analysis. We fit three ordinal regression models:
\begin{itemize}
    \item \textbf{Model A:} overcorrection $\sim$ probe $+$ threshold $+$ framing $+$ model
    \item \textbf{Model B:} Model A $+$ scenario fixed effects
    \item \textbf{Model C:} Leading-probe trials only: overcorrection $\sim$ threshold $+$ framing $+$ model $+$ scenario
\end{itemize}
All $p$-values are corrected for multiple comparisons using Benjamini-Hochberg FDR ($q < 0.05$).

\subsection{Study 2: Momentum Experiment}
\label{sec:momentum-methods}

Study~1 uses a fixed two-turn structure. A natural question is whether prior revision history shifts the revision gate on subsequent turns. Study~2 tests this with a multi-turn design in which the model first receives 1, 2, or 3 rounds of the leading probe (``Can this be improved?'') before encountering the evaluative probe (``Take another look at this and let me know if it's ready''). We call the number of prior leading-probe rounds the \textit{momentum dose}.

\paragraph{Design.} The momentum experiment crosses four factors: model (3) $\times$ threshold level (4: 0, 75, 90, 100) $\times$ threshold framing (2: numeric, qualitative) $\times$ momentum dose (3: 1, 2, or 3 prior leading-probe rounds) $\times$ scenario (8) $\times$ 3 runs $=$ \textbf{1,728 trials}. The no-momentum baseline (dose $= 0$) comes from Study~1's evaluative-probe trials, where the model received no prior ``Can this be improved?'' rounds before the evaluative probe.

\paragraph{Procedure.} Each trial proceeds as follows: the user sends the scenario prompt with an embedded threshold (Turn~1). For each dose $d$, the model responds, the user sends ``Can this be improved?'', and the model revises. After $d$ such rounds, the user sends the evaluative probe. The model's final response is scored by the same GPT-4o judge used in Study~1, using an adapted rubric that accounts for the multi-turn context.

\paragraph{Analysis.} We use chi-squared tests (RQ4: no-momentum vs.\ momentum gate distributions), Spearman correlation (RQ5: dose vs.\ revision rate), and logistic regression (RQ6: revised $\sim$ dose $+$ threshold $+$ dose $\times$ threshold) to test three research questions about how momentum interacts with the revision gate.

\paragraph{Reverse Momentum.} To test whether conversational framing can suppress revision below the no-momentum baseline, we add a reverse condition (dose $= -1$) in which the model receives one round of affirming feedback (``This looks great, no changes needed'') before the evaluative probe. The reverse condition uses the same scenario $\times$ threshold $\times$ framing $\times$ model design as Study~2, yielding 485 additional scored trials.

\paragraph{Inter-Rater Reliability.} A stratified sample of 60 momentum trials (balanced across model, threshold level, and dose) was re-scored by Claude Sonnet 4 as a second judge (Table~\ref{tab:irr-momentum}). Agreement on revision value is good ($\kappa = 0.669$) and revision magnitude is acceptable ($\kappa = 0.575$). Overcorrection ($\kappa = 0.392$) and threshold alignment ($\kappa = 0.339$) show lower agreement, partly attributable to the low variance in these scores under momentum (most trials are ``decline'' with scores of 1). Within-1 agreement exceeds 85\% on all dimensions.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Dimension} & \textbf{QW $\kappa$} & \textbf{\% Agree} & \textbf{\% Within 1} \\
\midrule
Revision magnitude & 0.575 & 68.3 & 85.0 \\
Revision value & 0.669 & 76.7 & 88.3 \\
Overcorrection & 0.392 & 71.7 & 90.0 \\
Threshold alignment & 0.339 & 38.3 & 85.0 \\
\bottomrule
\end{tabular}
\caption{Inter-rater reliability for Study~2 (quadratic-weighted Cohen's $\kappa$, GPT-4o vs.\ Claude Sonnet 4, $n = 60$).}
\label{tab:irr-momentum}
\end{table}
