\section{Related Work}

\paragraph{Sycophancy and Instruction Compliance.}
Instruction-tuned LLMs exhibit a well-documented tendency toward sycophantic behavior: agreeing with users, complying with requests, and avoiding disagreement even when it would be more helpful \cite{perez2023discovering, sharma2024towards, wei2024simple}. \citet{ranaldi2023large} show that this tendency extends to settings where models comply with implied requests even when the user's own stated preferences suggest otherwise. Our work documents a specific pattern consistent with this compliance tendency in revision tasks: a phrasing-sensitive revision gate that fires regardless of the user's stated quality standard.

\paragraph{RLHF and Alignment Training.}
The revision gate's bias toward compliance is plausibly shaped by reinforcement learning from human feedback (RLHF), in which reward models trained on human preference data may systematically reward helpful action over restraint \cite{ouyang2022training, bai2022training}. Direct preference optimization \cite{rafailov2023direct} inherits this bias to the extent that training pairs favor responses that ``do something'' over responses that decline. Our findings are consistent with this hypothesis: all three models, despite different training pipelines, exhibit near-universal compliance under revision-implying prompts, suggesting a shared structural incentive.

\paragraph{Prompt Sensitivity.}
LLMs are highly sensitive to surface-level features of prompts (phrasing, formatting, and ordering), often more than to semantic content \cite{lu2022fantastically, zhao2021calibrate, sclar2024quantifying}. \citet{mizrahi2024state} demonstrate that minor reformulations of equivalent instructions produce substantially different outputs. We extend this literature to the revision setting, showing that the pragmatic force of a follow-up prompt (revision-implying vs.\ evaluative) dominates a semantically meaningful constraint (the quality threshold). Notably, this is a multi-turn phenomenon: the threshold is stated in Turn~1 but the gate-controlling signal arrives in Turn~2, raising the possibility that recency weighting contributes to threshold insensitivity.

\paragraph{LLM Self-Refinement.}
A growing body of work explores LLMs' ability to iteratively refine their own outputs. \citet{madaan2024self} propose Self-Refine, showing that models can improve outputs through prompted self-critique. \citet{huang2024large} argue that models cannot reliably self-correct reasoning without external feedback. \citet{kim2024language} examine conditions under which self-refinement succeeds or fails in reasoning tasks. These studies typically assume that refinement is desirable; we examine the complementary question of whether models can \textit{refrain} from refinement when the user's quality standard has already been met.

\paragraph{LLM-as-Judge.}
Using LLMs to evaluate generated text has become standard practice \cite{zheng2024judging, li2024generation}. Concerns about self-preferencing bias, where a model favors its own outputs, are well-documented \cite{panickssery2024llm, koo2024benchmarking}. We address this by testing for judge bias directly, using a second judge model for inter-rater reliability, and employing a judge-independent response length proxy that is broadly consistent with judge rankings.

\paragraph{Controllable Generation and Quality Calibration.}
Holistic evaluation efforts such as HELM \cite{liang2023holistic} have assessed model capabilities across style, toxicity, and formality dimensions, but the question of whether models calibrate \textit{revision behavior} to user-stated quality thresholds has received limited direct attention. \citet{zhou2024calibrated} study calibration in the context of next-token prediction, and \citet{lin2022teaching} explore models' ability to express uncertainty. Our work is, to our knowledge, the first to empirically test whether the decision to revise (not just the quality of the output) is sensitive to user-stated standards.
