% ACL 2024 Paper
% Download acl.sty and acl_natbib.bst from:
%   https://github.com/acl-org/acl-style-files-and-templates
% Place them in this directory before compiling.

\documentclass[11pt]{article}

% --- ACL style (comment out the fallback block below when using real acl.sty) ---
\usepackage[hyperref]{acl}

\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{xcolor}
% Brand colors matching matplotlib figures
\definecolor{claudeorange}{HTML}{D97757}
\definecolor{googleblue}{HTML}{4285F4}
\definecolor{gptdark}{HTML}{1A1A2E}
\definecolor{gategreen}{HTML}{2ECC71}
\definecolor{gateamber}{HTML}{F39C12}
\definecolor{gatered}{HTML}{E74C3C}
\usepackage{tabularx}
\usepackage{tikz}
\usetikzlibrary{positioning,calc,decorations.pathreplacing}

\title{When LLMs Revise Beyond the User-Defined Threshold: \\
Prompt Phrasing Controls the Revision Gate}

\author{
  Liam Neild \\
  Emory University \\
  \texttt{liam.neild@emory.edu}
}

\begin{document}
\maketitle

\input{sections/abstract}
\input{sections/introduction}
\input{sections/related_work}
\input{sections/methods}
\input{sections/results}
\input{sections/discussion}
\input{sections/conclusion}

\section*{Limitations}

We do not control or evaluate the quality of Turn~1 outputs. If models produce outputs that already exceed the stated threshold, then revision may be rational regardless of the threshold level. A direct quality assessment of Turn~1 outputs would strengthen the claim that models are revising unnecessarily.

Our main experiment uses two probes selected from a five-probe pilot. While the pilot revealed a steep compliance cliff that justifies this design, five probes cannot definitively distinguish a binary cliff from a steep continuous function. A finer-grained probe gradient (e.g., 15--20 wordings spanning the revision-evaluation spectrum) would better characterize the transition region and determine whether intermediate compliance rates exist.

Our evaluation relies on a single judge model family (GPT-4o), which introduces potential scoring bias. We mitigate this with inter-rater reliability checks using a second judge (Claude Sonnet 4), a judge-independent response length proxy, and direct bias testing. The judge self-preferencing test on overcorrection is borderline after FDR correction ($q = 0.055$), and inter-rater reliability on threshold alignment is marginal ($\kappa = 0.556$).

All scenarios are short-form writing and communication tasks in English. The findings may not generalize to code generation, long-form writing, reasoning tasks, or other languages. Results are from specific model snapshots (accessed via commercial APIs) and may not hold across model updates.

The design uses 5 runs per cell. Within-cell variance is low (ICC(1) $= 0.71$), confirming near-deterministic behavior at low temperature. The conservative effective sample size is 768 unique cells rather than the nominal $N = 3{,}840$. All trials use a fixed two-turn structure; real conversations involve longer histories where thresholds might accumulate force across turns.

\section*{Ethics Statement}

This study involves no human subjects. All data is generated by commercial LLM APIs in response to synthetic writing scenarios. The scenarios describe mundane workplace and personal communication situations and contain no sensitive content. We do not release model outputs that could be mistaken for human-written text.

\bibliography{references}

\appendix
\input{sections/appendix}

\end{document}
